---
title: "model"
author: "Antoneely Calizaya"
date: "2/12/2022"
output: pdf_document
---

```{r}
#install packages
install.packages("tidyverse")
install.packages("tidymodels")
install.packages("scales")
install.packages("tinytex")
install.packages("googledrive")
install.packages("glmnet")

library(tidyverse)
library(tidymodels)
library(scales)
library(tinytex)
library(googledrive)
library(glmnet)

```

```{r}
#import the dataset from googledrive
drive_auth()

drive_find(n_max = 30)

filep <- drive_download("data.csv", overwrite = F)
filep$path
filep$local_path

#dataframe formed from the csv file

carsales <- read.csv("data.csv", header = TRUE, sep = ",")
sdata
````

```{r}
#Frequency Polygon - log scale to see the data distribution
carsales %>%
  ggplot(aes(price))+
  geom_freqpoly()

```
```{r}
#Frequency Polygon - with a log scale to see the data distribution
carsales %>%
  ggplot(aes(price))+
  geom_freqpoly() +
  scale_x_log10(label = label_dollar()) +
  theme_minimal()
```
```{r}
#Boxplots allow us to see the distribution of the individual variables to see how they contribute to the result ‘price’.
carsales %>%
  ggplot(aes(x= manufacturer, y = price)) +
  scale_y_log10(labels = label_dollar())+
  geom_boxplot() +
  coord_flip()

carsales %>%
  ggplot(aes(x=condition, y=price, fill = transmission))+
  geom_bar(position="dodge", stat="identity")+
  scale_y_log10(label = label_dollar())+
  theme_minimal()

```
```{r}
#year wheb the car was built
#linear method
ggplot(carsales, aes(x= year, y=price)) +
  geom_point() +
  scale_y_log10(labels = label_dollar())+
  geom_smooth(method = lm) + theme_minimal() +
  labs( x = "Year Vehicle Was Built", y = "Listed Price" )

```
```{r}
#removing linear method
ggplot(carsales, aes(x= year, y=price)) +
  geom_point() +
  scale_y_log10(labels = label_dollar())+
  geom_smooth() + theme_minimal() +
  labs( x = "Year Vehicle Was Built", y = "Listed Price" )

```
The linear relationship is evident for the cars built after the year 2000 - consider limiting the model to these cars only

## Grouping many groups into specific categories
The vehicles in the dataset are classified into 13 different types, each with four size groups.
```{r}
#grouping groups into specifics categories
ggplot(carsales, aes(x = size, fill = type)) +
  geom_bar() + facet_wrap(vars(type))+
  theme(axis.text.x = element_text(size = 8, angle = 45)) +
  theme_minimal() +labs( x = "Size of Listed Vehicle")

```
```{r}
carsales <- carsales %>%
   transmute(type = case_when(type == "truck" | type == "pickup" | type == "offroad" ~ "truck", 
                              type == "sedan" ~ "sedan",
                              type == "SUV" ~ "SUV",
                              type == "van" | type == "mini-van" | type == "bus" ~ "van",
                              TRUE ~ "other"), year, manufacturer ,model, condition, cylinders, transmission, fuel, title_status, size, odometer,region, lat, long, price)

str(carsales)
```

```{r}
carsales_2000 <- carsales %>% filter(year >= 2000)

mean(carsales$price)
mean(carsales_2000$price)

```
# Feature Engineering
```{r}
ggplot(carsales_2000, aes(x = size, fill = type)) +
  geom_bar() + facet_wrap(vars(type))+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  theme_minimal() +labs( x = "Size of Listed Vehicle")
```
```{r}

carsales_2000 %>% count(region, sort = TRUE)

```
# Develop the regression models using the tidymodels packages
```{r}

# rename the carsales df to something smaller

cars2k <- carsales_2000
# remove the two other dfs to free up memory 

rm(carsales)
rm(carsales_2000)
```

```{r}
#set seed to anchor the randomness to allow reproducibility in the data splitting process - otherwise you will get a different split each time you run  
set.seed(54321)

# data split

cars2k_split <- initial_split(cars2k)
train_data <- training(cars2k_split)
test_data <- testing(cars2k_split)

```

# Set the linear regression recipe

```{r}

recipe_lm <- recipe(price ~  manufacturer+condition + type+size+transmission+fuel, data=train_data)%>%
  step_dummy(all_nominal_predictors())%>%
  step_center(all_numeric_predictors())%>%
  step_scale(all_numeric_predictors())

prep_rec<-  prep(recipe_lm, data = train_data)

```

```{r}
# set the model
model_lm <- linear_reg(penalty = double(1))%>%
  set_engine("glmnet")
```

```{r}
# Set the workflow

wf_lm <- workflow() %>%
  add_recipe(recipe_lm)%>%
  add_model(model_lm)

```

```{r}
# fit the model

lm_fit <- wf_lm %>%
  fit(data = train_data)

lm_fit
```
```{r}
# predict the result on test data using augment function

final_data <- augment(lm_fit, new_data = test_data)%>%
  select(price, type, .pred)%>%
  mutate(variance = round(price-.pred))

```

```{r}

#plot

final_data %>%
  ggplot(aes(x=.pred, y=type))+
  geom_boxplot()+
  scale_x_log10(labels = label_dollar())+
  theme_minimal() +
  labs( x = "Predicted price", y = "Type of vehicle", title = "Predicted price of cars based on its type (linear modelling)" )

 view(final_data)
```