---
title: "EDA"
author: "Antoneely Calizaya"
date: "21/02/2022"
output: pdf_document
---

```{r}
install.packages("tidyverse")
install.packages("skimr")
install.packages("janitor")
install.packages("ggplot2")
install.packages("scales")
install.packages("glue")
install.packages("tidyselect")
library(tidyverse)
library(skimr)
library(janitor)
library(ggplot2)
library(scales)
library(glue)
library(tidyselect)
```


```{r}
train_data <- 
  read_csv("train.csv",
           show_col_types = FALSE) %>%
  clean_names() %>%
  select(id, dplyr::everything())

skim(train_data)
```
1. Variables with missing values, replace with NA or mean, average (year_built,
energy_star_rating, direction_max_wind_speed,  direction_peak_wind_speed, max_wind_speed,days_with_fog)
2. Consider to agroup the Average temperature into 4 seasons or total year
3. Check correlation betweeten building_class, facility_tye with site_eui

```{r}
#convert conitnous variable to factor
train_data$year_factor <-as.factor(train_data$year_factor)

#replaced "0" values to NA in year_built variable
train_data <- train_data %>%
  mutate(year_built = na_if(year_built, 0))

```

Understanding the missing values

```{r}
train_data %>% skim() %>%
  filter(n_missing != 0) %>%
  as_tibble() %>%
  select(skim_variable, n_missing, complete_rate) %>%
  mutate(missing_rate = round(abs(complete_rate - 1) * 100, 1)) %>%
  ggplot(aes(x=fct_reorder(skim_variable, n_missing), y=missing_rate,fill = skim_variable,label = paste0(missing_rate, "%", " / ", n_missing)))+
  geom_col(width = .5) +
  theme_minimal()
  
```

```{r}
train_data %>% skim() %>%
  filter(n_missing != 0) %>%
  as_tibble() %>%
  select(skim_variable, n_missing, complete_rate) %>%
  mutate(missing_rate = round(abs(complete_rate - 1) * 100, 1)) %>%
  ggplot(aes(
    x = fct_reorder(skim_variable, n_missing),
    y = missing_rate,
    fill = skim_variable,
    label = paste0(missing_rate, "%", " / ", n_missing)
  )) +
  geom_col(width = .5) +
  geom_text(aes(
    hjust = case_when(n_missing == min(n_missing) ~ -0.1,
                      TRUE ~ 1.2),
    color = case_when(n_missing == min(n_missing) ~ "black",
                      TRUE ~ "white")
  )) +
  coord_flip() +
  theme(legend.position = "none") +
  scale_y_continuous(label = label_percent(scale = 1)) +
  scale_fill_manual(values = c(
    "#e41a1c",
    "#984ea3",
    "#4b2b31",
    "#090966",
    "#d6493c",
    "#203d51"
  )) +
  scale_color_manual(values = c("black", "white")) +
  labs(
    title = "Missing values within each variable",
    caption = "Data source: Kaggle.com, WiDS_Datathon_2022",
    x = "Variable name",
    y = "Percent"
  )

```
```{r}
eui_outlier_fun <- function(var_to) {
  
var_title <- as_label(enquo(var_to))
  
var_title <- str_to_title(str_replace(var_title, pattern = "_", replacement = " " ))

  train_data %>% 
  select(site_eui, {{var_to}} ) %>%
  ggplot(aes(x = as.factor({{var_to}}), 
             y = site_eui)) +
    geom_boxplot(
      fill = "#D5B488",
      color = "#27A118",
      alpha = 0.5,
      outlier.color = "#F6645A",
      size = 0.5
    ) +
    coord_flip() +
    theme(axis.text.y = element_text(size = 5),
          axis.ticks.y = element_blank()) +
    labs(
      title = glue("Site EUI outliers by {var_title}"),
      subtitle = NULL,
      x = glue("{var_title}"),
      y = NULL
    )
}

p1 <- eui_outlier_fun(state_factor) 

p2 <- eui_outlier_fun(building_class)

p3 <- eui_outlier_fun(year_factor)

p4 <- eui_outlier_fun(elevation)

```

```{r}
ggplot(train_data, aes(state_factor, site_eui, colour = state_factor, alpha = 0.2))+
  geom_point(alpha = 0.2)
```
```{r}
ggplot(train_data, aes(building_class, site_eui, alpha = 0.2))+
  geom_point(alpha = 0.2)
```
```{r}
ggplot(train_data, aes(year_factor, site_eui, colour = year_factor, alpha = 0.2))+
  geom_point(alpha = 0.2)
```

```{r}
ggplot(train_data, aes(energy_star_rating, site_eui, colour = year_factor, alpha = 0.2))+
  geom_point(alpha = 0.2)
```

```{r}
ggplot(train_data, aes(floor_area, site_eui, alpha = 0.2))+
  geom_point(alpha = 0.2)
```
Buildings with less square feet use more energy.

```{r}
ggplot(train_data, aes(floor_area, building_class, alpha = 0.2))+
  geom_point(alpha = 0.2)
```

ot of building more than 1e+6 square feet majority of them are commercial class! and the floor area outliers are mainly related to commercial buildings! it is a piece of important information!

```{r}
ggplot(train_data, aes(site_eui,facility_type, colour = year_factor))+
  geom_point(alpha = 0.1)+
  theme_minimal()
```